---
title: "Covid_NYS_bargraph"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(ggplot2)
library(ggthemes)
library(extrafont)
library(tidyverse)
library(scales)

```

```{r}
options(scipen = 999)
```


```{r}
p_and_c<- read_csv("Pop_And_Count.csv")
```



Lets add the days of the week to our Df
```{r}
p_and_c<- p_and_c %>%
mutate(day_week = wday(Test_Date,label = TRUE))
```

Now let us look at every Monday
```{r}
p_and_c<-filter(p_and_c, day_week == "Mon")
```


```{r}
View(p_and_c)
```

```{r figl, fig.height=5, fig.width=4}
B1<- ggplot(p_and_c,aes(x=p_and_c$County, y= p_and_c$Cumulative_Number_of_Positives))+
      geom_bar(stat = "identity")+
  coord_flip()+
  ggtitle("Unadjusted #s for Each county As of April 8,2020")
B1
```




```{r  figl, fig.height=5, fig.width=4}


```

```{r figl, fig.height=5, fig.width=4}
L1<- ggplot(p_and_c, aes(x=p_and_c$County  ,y=p_and_c$Cumulative_Number_of_Positives))+
      geom_line()+coord_flip()
L1
```


```{r figl, fig.height=5, fig.width=4}
L2<- ggplot(p_and_c, aes(x=p_and_c$County  ,y=p_and_c$Cumulative_Number_of_Positives))+
      geom_line(aes(width=1))+coord_flip()
L2
```


### We need to make the data skinny and out all of the ethnicities 
### under one variable in order to create a stacked bar plot.
### Convert COUNTY to a factor
####http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/
```{r}
ethnicities$County<- as.factor(ethnicities$County)
```

```{r}
dim(ethnicities)
```

```{r}
head(ethnicities,3)
```

```{r}
ethnicities_long<- gather(ethnicities , race , number, White:All_others )
```

```{r}
head(ethnicities_long)
```

```{r}
dim(ethnicities_long)
```


###stacked bar  http://t-redactyl.io/blog/2016/01/creating-plots-in-r-using-ggplot2-part-4-stacked-bar-plots.html

### We need  to mutate adn add a column for labels
```{r}
ethnicities_long_2 <- ethnicities_long%>%
                      mutate(pct = number/Total)
```

```{r}
head(ethnicities_long_2)
```

###Let us round that decimal
```{r}
ethnicities_long_2$pct <-  format(round(ethnicities_long_2$pct , 2) ,nsmall = 2 )
```

```{r}
ethnicities_long_2$pct <- as.numeric(ethnicities_long_2$pct)
```


```{r}
head(ethnicities_long_2,3)
```

```{r}
str(ethnicities_long_2)
```


```{r  figl, fig.height=5, fig.width=4}
b1<- ggplot(data = ethnicities_long_2, aes(x= County , y = number)) +
        geom_col(aes(fill= race) ,width = 0.7) +
            #geom_text(aes(y=pct, label = number, group=County) , color="black")+
            coord_flip()+
ggtitle("NYS County Pop")
b1
       
```


100% Stacked bar graph

```{r  figl, fig.height=5, fig.width=4}
b2<- ggplot()+
      geom_bar(aes(y=pct , x= County, fill=race) , data = ethnicities_long_2, stat = "identity")+
  coord_flip()
b2

```


#https://stackoverflow.com/questions/9563368/create-stacked-barplot-where-each-stack-is-scaled-to-sum-to-100


```{r figl, fig.height=5, fig.width=4 }
b3<- ggplot(ethnicities_long_2, aes( x= County, y= pct,fill= race))+
      geom_bar(position = "fill", stat="identity")+
      coord_flip()+
      ggtitle("Look At How While NYS Is")+
  theme_bw()
b3
```



```{r}
head(NYS_Covid_Data,3)
```


b1<- ggplot(data = ethnicities_long_2, aes(x= County , y = number)) +
        geom_col(aes(fill= race) ,width = 0.7) +
            #geom_text(aes(y=pct, label = number, group=County) , color="black")+
            coord_flip()+
ggtitle("NYS County Pop")
b1

```{r  figl, fig.height=5, fig.width=4}
c1<- ggplot(data = NYS_Covid_Data,aes(x= County , y = New_Positives)) +
      geom_col(  width = 0.7) +
  coord_flip()+
  ggtitle(" Number of Confirmed Covid Cases as of April 9, 2020 ")+
  theme_economist()

c1
```







