---
title: "EuroBarplot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Preferences
#Turn off scientific notation
```{r}
options(scipen = 999)
```

#Packages
```{r}
library(tidyverse)
library(ggplot2)
```

import data
make skinny
bar plot group by coountry
maybe 2 axis

```{r}
```


```{r}
data<- read_csv("country_pop_covid.csv")
data
```

```{r}
head(data,2)
```

```{r}
data$Country<- as.factor(data$Country)
```

bar plot side by side grouped by country and stacked bar.
make data skinny
 
#ethnicities_long<- gather(ethnicities , race , number, White:All_others )


```{r}
head(data,2)
```



```{r}
A<- ggplot(data, aes(x=Country, y= Population))+
  geom_bar(stat = "identity")+
  coord_flip()
A
```

# stacked, grouped, dodge, multiple groups

```{r}
B<- ggplot(data, aes(x=Country, y= Population))+
    geom_col(aes(color= "black", fill=""))
```










#############
##############3
```{r}
#country_pop_long<- gather(data, incidences,number, Covid_deaths:deaths_per_100k )
```

```{r}

```

```{r}
head(country_pop_long)
```

```{r}
str(country_pop_long)
```

```{r}
View(country_pop_long)
```
```{r}
str(country_pop_long)
```

b1<- ggplot(data = ethnicities_long_2, aes(x= County , y = number)) +
        geom_col(aes(fill= race) ,width = 0.7) +
            #geom_text(aes(y=pct, label = number, group=County) , color="black")+
            coord_flip()+
ggtitle("NYS County Pop")
b1



```{r}
A<- ggplot(country_pop_long, aes(x=Country, y= Population))+
  geom_col(aes(fill=incidences), position = position_dodge(0.8), width = 0.7)+
  coord_flip()
A
```


```{r}
df
```

```{r}
f<- ggplot(df, aes(x-dose, y=len))
```


```{r}
f <- ggplot(df, aes(x = dose, y = len))
```

```{r}
f + geom_col(fill = "#0073C2FF")+
  geom_text(aes(label = len), vjust = 1.6, color = "white")
```










 
