---
title: "SK_US_tests_v_deaths"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Link to GoogDrive did not work.
```{r}
library(googledrive)
```
#https://googledrive.tidyverse.org/

```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)

```

```{r}
options(scipen = 999)
```


```{r}
#drive_find( pattern = "covid-tests-cases-deaths")
```

# data  https://drive.google.com/open?id=1AWZ8TKpnka7Mn3kjHSo3Rjf9Su3RTSJY

#sites that show combo plots
#This one shows 2 lines and side by side plots:
#https://stackoverflow.com/questions/44406921/how-to-combine-line-and-bar-chart-in-ggplot2-with-just-one-grouping-variable

#https://stackoverflow.com/questions/47940161/ggplot-plotting-bars-and-lines-in-the-same-chart

#https://stackoverflow.com/questions/41764312/combining-bar-and-line-chart-double-axis-in-ggplot2


'''
The plan.
Create a plot that has a line for US and a seperate line for SK that shows the monthly  number of tests starting from whatever date both got first confirmed case.

These two lines will be on the same graph as a side by side bar plot that shows the monthly death total PER 100k for each country.
Because the data is such a mess for both he test data and the death data I will seperate US & Sk and work on each of the 4 parts indivudally.


Need to make rename cols, drop extraneous cols, make country categorical, make date date type, add mutate in country pop, mutate in tests / deaths per 100k.

After all of that rbind US&SK back together , so where I had 4 now I have two data sets .

      * rename columns
      *drop extraneous columns
      #concatenate 1st & 2nd columns
      *change COUNTRY to factor
      *change DATE to date-type
      * filter out extraneous dates
      *add column of population
      * add column of tests per 100k
      
'''

```{r}
data<- read_csv("covid_tests.csv")
```
                   ######################
##########################  US TESTS ##################3
                   ########################
# First Test Data
```{r}
View(data)
```

```{r}
# First we will filter out the US test data 
#There may not be a single column in this Data properly named.
US_tests<- filter(data,Date == "USA")
```

```{r}
#rename cols
#rename(new name = existing)
#US_deaths<- US_deaths %>%
            #rename(Country = Entity ,Deaths = `Daily confirmed deaths (deaths)`)
US_tests<- US_tests %>%
            rename(Country = Entity ,Second= Code, Code = Date, Date = Cumulative)
```

```{r}
#drop extraneous cols
## Notice the lack of quotes
#iris %>% select (-c(Sepal.Length, Sepal.Width))
## Notice the lack of quotes
US_tests<-US_tests %>% select (-c(Total, confirmed, deaths, (deaths), Total_1,
                        confirmed_1, cases, (cases)))
```

```{r}
# variables names in parathesis not dropped above
#df$x <- NULL
US_tests$`(deaths)`<- NULL
```

```{r}
# variables names in parathesis not dropped above
#df$x <- NULL
US_tests$`(cases)`<- NULL
```


```{r}
# 'UNITED' and 'STATES' are in seperate cols.  Concatenate.
US_tests$Country<- paste(US_tests$Country, US_tests$Second)
```

```{r}
#Now "SECOND" is extraneous
US_tests$Second <- NULL
```

```{r}
US_tests$Country<- as.factor(US_tests$Country)
```

```{r}
US_tests$Date <- mdy(US_tests$Date)
```

```{r}
#US & SK each got the first confirmed case on thesame day
#We will filter out dates BEFORE that

US_tests <-US_tests %>%
        filter(Date >= as.Date("2020-01-20") & Date <= as.Date("2020-04-20"))
```

```{r}
#We will calcculate teste per 100k but first we need to add_case(collumn filled with #US population .  328,200,000)
# main$newcol <- rep(1,nrow(main))
US_tests$population <-rep(328200000, nrow(US_tests))
```

```{r}
#Some of our "test" data is NA.  
#We need to change those to 0's
US_tests$tests[is.na(US_tests$tests)] <-0
```

```{r}
#make sure TESTS are numeric
US_tests$tests <- as.numeric(US_tests$tests)
```

```{r}
# Using MUTATE we generate new column based on existing
US_tests<- US_tests %>%
    mutate(per_100k = (tests / population)*100000)
```


```{r}
US_tests[ , 'per_100k']= round(US_tests[ , 'per_100k'] , 1)
```



```{r}
str(US_tests)
```
First confirmation January 20, 2020



```{r}
View(US_tests)
```
                                 #########################3
                    ################## Now SK tests ##################
                               ##############################

```{r}
# First we will filter out the US test data 
#There may not be a single column in this Data properly named.
SK_tests<- filter(data,Code == "Korea")
```


```{r}
#rename cols
#rename(new name = existing)
#US_deaths<- US_deaths %>%
            #rename(Country = Entity ,Deaths = `Daily confirmed deaths (deaths)`)
SK_tests<- SK_tests %>%
            rename(Country = Entity ,Second= Code, Code = Date, Date = Cumulative)
```


```{r}
#drop extraneous cols
## Notice the lack of quotes
#iris %>% select (-c(Sepal.Length, Sepal.Width))
## Notice the lack of quotes
SK_tests<-SK_tests %>% select (-c(Total, confirmed, deaths, (deaths), Total_1,
                        confirmed_1, cases, (cases)))
```

```{r}
SK_tests$`(deaths)`<- NULL
```

```{r}
SK_tests$`(cases)`<- NULL
```


```{r}
SK_tests$Country<- paste(SK_tests$Country, SK_tests$Second)
```


```{r}
SK_tests$Second<- NULL
```

```{r}
SK_tests$Country<- as.factor(SK_tests$Country)
```

```{r}
SK_tests$Date <- mdy(SK_tests$Date)
```

```{r}
SK_tests <-SK_tests%>%
          filter(Date >= as.Date("2020-01-20") & Date <= as.Date("2020-04-20"))
```


```{r}
SK_tests$population <- rep(51640000, nrow(SK_tests))
```

```{r}
#We need to change those to 0's
SK_tests$tests[is.na(SK_tests$tests)] <-0
```


```{r}
SK_tests$tests <- as.numeric(SK_tests$tests)
```

```{r}
SK_tests<- SK_tests %>%
    mutate(per_100k = (tests / population)*100000)
```

```{r}
SK_tests[ , 'per_100k']= round(SK_tests[ , 'per_100k'] , 1)
```


```{r}
View(SK_tests)
```

```{r}
str(SK_tests)
```



                         #############################             
               ############ NOW JOIN US & SK TESTS  ####################
                         ##############################         
               
               
               
```{r}
Tests_rbind<- bind_rows(US_tests , SK_tests)
```

```{r}
inner<- inner_join(US_tests, SK_tests ,by= "Date")
```



#https://www.storybench.org/pivoting-data-from-columns-to-rows-and-back-in-the-tidyverse/

# NOTE pivot longer requires all data to be of same data type, cannot mix factors and numeric and data , etc





#################### PLOT TESTS ####################

```{r}
ggplot(inner, aes(Date)) +
  geom_line(aes(y=per_100k.x) , color= "red")+
  geom_line(aes(y=per_100k.y), color= "blue")
```


```{r}
ggplot(inner, aes(Date)) +
  geom_line(aes(y=per_100k.x,color=per_100k.x))+
  geom_line(aes(y=per_100k.y, color=per_100k.y))
```


  ############################################################################################
              #####    Now wrangle Us and SK death data                                  ######
#################################################################################
* rename columns
      *Filter to get just the single country
      *drop extraneous columns
      *rename cols
      #concatenate 1st & 2nd columns
      *change COUNTRY to factor
      *change DATE to date-type
      * filter out extraneous dates
      * add column of population
      * add column of tests per 100k


```{r}
covid_deaths<- read_csv("covid_daily_deaths.csv")
```


```{r}
View(covid_deaths)
```

```{r}
US_Deaths<- filter(covid_deaths,Code == "USA")
```



```{r}
#Change col names
#rename(new name = existing)
US_Deaths<- US_deaths %>%
            rename(Country = Entity ,Deaths = `Daily confirmed deaths (deaths)`)
```

```{r}
US_Deaths$Country<- as.factor(US_Deaths$Country)
```


```{r}
US_Deaths$Date <- mdy(US_Deaths$Date)
```

```{r}
US_Deaths$Population <- rep(328200000, nrow(US_Deaths))
```


```{r}
US_Deaths<- US_Deaths %>%
          mutate(per_100k = (Deaths / Population)* 100000)
```

```{r}
US_Deaths <- US_Deaths%>%
          filter(Date >= as.Date("2020-01-20") & Date <= as.Date("2020-04-20"))
```


```{r}
B<- US_Deaths  %>%
    mutate(month = format(Date, "%m"))  %>%
                          group_by(month)  %>%
                            summarise(total = sum(per_100k))
```

```{r}
B
```



```{r}
B$month<- as.numeric(B$month)
```

```{r}
US_Deaths_monthly <- B %>% mutate(MonthName = month.name[month])
```

```{r}
str(US_Deaths_monthly)
```

```{r}
#US_Deaths_monthly$MonthName<- as.numeric(US_Deaths_monthly$MonthName)
```

```{r}
US_Deaths_monthly

```

```{r}
US_Deaths_monthly
```



```{r}
ggplot(US_Deaths_monthly,aes(x= MonthName, y= total))+
  geom_bar(stat = "identity")
```

```{r}
US_Deaths_monthly$MonthName<- factor(US_Deaths_monthly$MonthName, levels = c("January", "February","March","April",
                                                                             "May","June", "July","August","September",
                                                                             "October", "November","December"))
```


```{r}
ggplot(US_Deaths_monthly ,aes(x= MonthName, y= total))+
  geom_col()
```




```{r}
A<- US_Deaths%>% 
    group_by(month=floor_date(Date,"30 days"))%>%
    summarize(per_100k = sum(per_100k))
A
```

```{r}
A
```



```{r}
View(US_Deaths)
```


```{r}
str(US_Deaths)
```




############### Now South Korean Deaths #############

* rename columns
      *Filter to get just the single country
      *drop extraneous columns
      *rename cols
      #concatenate 1st & 2nd columns
      *change COUNTRY to factor
      *change DATE to date-type
      * filter out extraneous dates
      * add column of population
      * add column of tests per 100k
      


```{r}
SK_Deaths<- filter(covid_deaths,Code == "KOR")
```



```{r}
SK_Deaths<- SK_Deaths %>%
            rename(Country = Entity ,Deaths = `Daily confirmed deaths (deaths)`)
```

```{r}
SK_Deaths$Date <- mdy(SK_Deaths$Date)
```

```{r}
SK_Deaths$Country <- as.factor(SK_Deaths$Country)
```


```{r}
SK_Deaths <-SK_Deaths%>%
          filter(Date >= as.Date("2020-01-20") & Date <= as.Date("2020-04-20"))
```


```{r}
SK_Deaths$population <- rep(51640000, nrow(SK_Deaths))
```

```{r}
SK_Deaths<- SK_Deaths %>%
    mutate(per_100k = (Deaths / population)*100000)
```


```{r}
View(SK_Deaths)
```

```{r}
str(SK_Deaths)
```

US_SK_Deaths_Wide


theme(axis.text.x = element_text(angle = 45, vjust = 0.75,hjust = 0.9))+
```{r  figl, fig.height=8, fig.width=4}
ggplot(SK_Deaths,aes(x= Date, y= SK_Deaths$per_100k, group =1))+
  geom_line()+
  coord_flip()
```



#####################################
# JOIN US & SK Deaths into wide

```{r}
dim(US_Deaths)  #92  6
```
```{r}
dim(SK_Deaths)  #113   6
```


```{r}
US_SK_Deaths_Wide<- bind_cols(US_Deaths, SK_Deaths)
```


```{r}
US_SK_Deaths_Narrow<- bind_rows(US_Deaths, SK_Deaths)
```

```{r}
View(US_SK_Deaths_Wide)
```

```{r}
View(US_SK_Deaths_Narrow)
```

```{r}
ggplot(US_SK_Deaths_Wide,aes(x=Date))+
  geom_line(aes(y=per_100k) , color = "red") +
  geom_line(aes(y=per_100k1) , color = "blue") +
  scale_colour_manual(values=c("red", "blue"))+
  ggtitle("Deaths Per 100,000 US (red) and SKorea(Blue")
```


```{r}
ggplot(US_SK_Deaths_Narrow, aes(x= Date, y= per_100k))+
      geom_col(aes(color= "black" ), fill= "green", position = position_dodge(0.9) , width = 0.9) +
      scale_color_manual(values = c("red", "blue"))+
      scale_y_log10()
```








```{r}
str(US_SK_Deaths_Narrow)
```

```{r}
US_SK_Deaths_Wide %>%
      pivot_longer(~Country ,)
```

```{r}
View(US_SKD)
```




















































































