---
title: "Corona_Countries_Pop_Adj"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
options(scipen = 999)
```

```{r}
library(ggplot2)
library(tidyverse)
library(gsheet)
library(choroplethr)
library(maps)
library(sf)
library("rnaturalearth")
library("rnaturalearthdata")
#
library(rworldmap)
#
require(maps)
require(ggmap)
library(maps)

```
# DATA NOVIA
#https://www.datanovia.com/en/blog/how-to-create-a-map-using-ggplot2/

```{r}
library(ggplot2)
library(dplyr)
require(maps)
require(viridis)
theme_set(
  theme_void()
  )
```

######## Monday, April 13,2020
# https://www.datanovia.com/en/blog/how-to-create-a-map-using-ggplot2/
* get world map data.
* make a list of countries of interst
* Subset world map data with western EU in our case.
* Join our data to subsetted world data.

```{r}
world_map <- map_data("world")

```


```{r}

ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill="lightgray", colour = "white")
```

#Make list of countries of interest
```{r}
# Some EU Contries
some.eu.countries <- c(
  "Portugal", "Spain", "France", "Switzerland", "Germany",
  "Austria", "Belgium", "UK", "Netherlands",
   "Italy", "luxembourg"
)
```

#Next we filter world map data by our selection
```{r}
# Retrieve the map data
some.eu.maps <- map_data("world", region = some.eu.countries)
```

```{r}
# Compute the centroid as the mean longitude and lattitude
# Used as label coordinate for country's names
region.lab.data <- some.eu.maps %>%
  group_by(region) %>%
  summarise(long = mean(long), lat = mean(lat))
```


```{r}
#plot given data
ggplot(some.eu.maps, aes(x = long, y = lat)) +
  geom_polygon(aes( group = group, fill = region))+
  geom_text(aes(label = region), data = region.lab.data,  size = 3, hjust = 0.5)+
  scale_fill_viridis_d()+
  theme_void()+
  theme(legend.position = "none")
```

#Works.  
#Now import our data
```{r}
data<- read_csv("country_pop_covid_deaths_3.csv")
```
```{r}
head(data,2)
```


#Next we join our data to the map data 
#But first we need to change covid col name from Country to "region"
```{r}
# rename( new name = existing name)
data<- data %>% 
                rename(region = Country)
```

```{r}
head(data,2)
```


```{r}
some.eu.maps<- full_join(some.eu.maps ,data, by = "region")
```


```{r}
some.eu.maps$deaths_per_100k<- as.factor(some.eu.maps$deaths_per_100k)
```

```{r}
ggplot(some.eu.maps, aes(x = long, y = lat)) +
  geom_polygon(aes( group = group, fill = some.eu.maps$deaths_per_100k))+
  geom_text(aes(label = region), data = region.lab.data,  size = 3, hjust = 0.5)+
  scale_fill_viridis_d()+
  theme_void()+
  theme(legend.position = "none")
```


```{r}
str(data_2)
```

```{r}
View(data_2)
```


'data.frame':	6021 obs. of  6 variables:
 $ long     : num  17 16.9 16.9 16.9 16.9 ...
 $ lat      : num  48.6 48.6 48.6 48.5 48.4 ...
 $ group    : num  1 1 1 1 1 1 1 1 1 1 ...
 $ order    : int  1 2 3 4 5 6 7 8 9 10 ...
 $ region   : chr  "Austria" "Austria" "Austria" "Austria" ...
 $ subregion: chr  NA NA NA NA ...
```{r}
str(data_2)
```
 



#Place labels
```{r}
# Compute the centroid as the mean longitude and lattitude
# Used as label coordinate for country's names
region.lab.data <- some.eu.maps %>%
  group_by(region) %>%
  summarise(long = mean(long), lat = mean(lat))
```

#Visualize







# Data
```{r}
data<- read_csv("country_pop_covid_deaths_3.csv")
```

#We need to join our covid data to the map dataset.
#First step rename covid "country" to region.

```{r}
# rename( new name = existing name)
country_pop_covid_deaths_3<- country_pop_covid_deaths_3 %>% 
                rename(region = Country)
```


```{r}
head(some.eu.maps,3)
```



```{r}
View(C)
```


```{r}
# Some EU Contries
some.eu.countries <- c(
  "Portugal", "Spain", "France", "Switzerland", "Germany",
  "Austria", "Belgium", "UK", "Netherlands",
  "Denmark", "Poland", "Italy", 
  "Croatia", "Slovenia", "Hungary", "Slovakia",
  "Czech republic"
)
```



```{r}
# Retrievethe map data
some.eu.maps <- map_data("world", region = some.eu.countries)
```



```{r}
View(some.eu.maps)
```

```{r}
dim(some.eu.maps)
```



```{r}
# Compute the centroid as the mean longitude and lattitude
# Used as label coordinate for country's names
region.lab.data <- some.eu.maps %>%
  group_by(region) %>%
  summarise(long = mean(long), lat = mean(lat))
```

```{r}
str(some.eu.maps)
```


ggplot(some.eu.maps, aes(x = long, y = lat)) +
  geom_polygon(aes( group = group, fill = region))+
  geom_text(aes(label = region), data = region.lab.data,  size = 3, hjust = 0.5)+
  scale_fill_viridis_d()+
  theme_void()+
  theme(legend.position = "none")
```{r   figl, fig.height=2, fig.width=2}
ggplot(some.eu.maps, aes(x = long, y = lat)) +
  geom_polygon(aes( group = group, fill = region))+
  geom_text(aes(label = region), data = region.lab.data,  size = 3, hjust = 0.5)+
  scale_fill_viridis_d()+
  theme_void()+
  theme(legend.position = "none")
```

```{r}
head(region.lab.data)
```








#########
########
```{r}
world_map <- map_data("world")
```

```{r}
ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill="lightgray", colour = "white")
```

#Map for specific regions


# Some EU Contries


```{r}
# Some EU Contries
some.eu.countries <- c(
  "Portugal", "Spain", "France", "Switzerland", "Germany",
  "Austria", "Belgium", "UK", "Netherlands",
  "Denmark", "Poland", "Italy"
)
```

```{r}
# Retrievethe map data
some.eu.maps <- map_data("world", region = some.eu.countries)
```

```{r}
# Compute the centroid as the mean longitude and lattitude
# Used as label coordinate for country's names
region.lab.data <- some.eu.maps %>%
  group_by(region) %>%
  summarise(long = mean(long), lat = mean(lat))
```




```{r   figl, fig.height=2, fig.width=2}
ggplot(some.eu.maps, aes(x = long, y = lat)) +
  geom_polygon(aes( group = group, fill = region))+
  geom_text(aes(label = region), data = region.lab.data,  size = 3, hjust = 0.5)+
  scale_fill_viridis_d()+
  theme_void()+
  theme(legend.position = "none")
```

# another source of mapping to address the distortions # https://timogrossenbacher.ch/2016/12/beautiful-thematic-maps-with-ggplot2-only/

#Make a choropleth Map
#World map colored by life expectancy

```{r}
library("WHO")
library("dplyr")
```

#https://github.com/expersso/WHO

#https://github.com/expersso/WHO
```{r}
library(devtools)
install_github("expersso/WHO")
```










########3
```{r}
help("country_choropleth")
```


```{r}
country_pop<- read_csv("country_pop.csv")
```

```{r}
str(country_pop)
```

```{r}
par(mfrow = c(2,1))
```

```{r}
map("usa")
```

```{r}
map("europe")
```
 #https://medium.com/@jules.beley/making-a-map-with-eu-data-on-r-erasmus-exchanges-by-country-3f5734dcd4ff
 
 
#############
# https://medium.com/@jules.beley/making-a-map-with-eu-data-on-r-erasmus-exchange#s-by-country-3f5734dcd4ff

```{r}
world<-map_data("world ")
```

```{r}
library("WHO")
library("dplyr")
```


```{r}
who
```

```{r}
life.exp <- get_data("WHOSIS_000001")             # Retrieve the data
```

```{r}
life.exp <- life.exp %>%
  filter(year == 2015 & sex == "Both sexes") %>%  # Keep data for 2015 and for both sex
  select(country, value) %>%                      # Select the two columns of interest
  rename(region = country, lifeExp = value) %>%   # Rename columns
  # Replace "United States of America" by USA in the region column
  mutate(
    region = ifelse(region == "United States of America", "USA", region)
    )                
```


```{r}
world_map <- map_data("world")
```

```{r}
life.exp.map <- left_join(life.exp, world_map, by = "region")
```

```{r}
ggplot(life.exp.map, aes(long, lat, group = group))+
  geom_polygon(aes(fill = lifeExp ), color = "white")+
  scale_fill_viridis_c(option = "C")
```

```{r}
ggplot(life.exp.map, aes(map_id = region, fill = lifeExp))+
  geom_map(map = life.exp.map,  color = "white", size=0.01)+
  expand_limits(x = life.exp.map$long, y = life.exp.map$lat)+
  scale_fill_viridis_c(option = "C")
```










